---
title: "Day Thirty-Four: Functions"
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
  `r sds::img_logo(width = 64)`
output:
  sds::moon_reader:
    lib_dir: libs
    css: xaringan-custom.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false    
      navigation:
        scroll: false
---

# Why write functions?

* Sometimes we find ourselves writing very similar lines of code over and over again for different data frames or variables. 
* Example:

```{r eval=FALSE}
df1 %>%
  summmarize(Total = sum(num_votes))

df2 %>%
  summmarize(Total = sum(num_votes))

df3 %>%
  summmarize(Total = sum(num_votes))
```

---

# Functions

* Statements organized to perform a specific task
* Take arguments as *inputs*
  * Arguments can be required or optional
* Return a value or set of values as *outputs*
  * Default value is output of last line of function if not specified

---

# User-defined Functions

Basic format:

```{r eval=FALSE}
function_name <- function(arg1, arg2) {

  x <- #Some code goes here referencing arg1 and arg2
  
  return(x)
}
```

---

# Calling Functions

> This is the same way you'd call functions built-in to `R`

```{r eval = FALSE}
function_name(value_for_arg1, value_for_arg2)
```

---

# User-defined Functions

```{r eval = FALSE}
sum_x_in_df <- function(data) {

  summed_df <- data %>%
    summarize(Total = sum(num_votes))
  
  return(summed_df)
}
```

```{r eval = FALSE}
sum_x_in_df(df1)
```

> `data` is a required argument above. 

---

# Making arguments optional

```{r eval = FALSE}
sum_x_in_df <- function(data, var = num_votes) {

  summed_df <- data %>%
    summarize(Total = sum({{var}}))
  
  return(summed_df)
}
```

```{r eval = FALSE}
sum_x_in_df(df1)
```

> Default value provided for `var`. 
> Use `{{...}}` to pass a variable name in a function. 

---

# Overriding Defaults

```{r eval = FALSE}
sum_x_in_df <- function(data, var = num_votes) {

  summed_df <- data %>%
    summarize(Total = sum({{var}}))
  
  return(summed_df)
}
```

```{r eval = FALSE}
sum_x_in_df(df1, num_registrations)
```

> `num_registrations` overrides the default value of `num_votes` for argument `var`

---

# Naming Arguments

```{r eval = FALSE}
sum_x_in_df <- function(data, var = num_votes) {

  summed_df <- data %>%
    summarize(Total = sum({{var}}))
  
  return(summed_df)
}
```

```{r eval = FALSE}
sum_x_in_df(data = df1, var = num_registrations)
```

> Order matters if arguments not named!

