---
title: "Day Four: Data Fundamentals and Intro to RStudio Environment"
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
output: pdf_document
---

```{r}
global_landslide <- read.csv("https://data.nasa.gov/api/views/dd9e-wu2v/rows.csv")
```

1. Identify a unique key in this dataset. Check whether this unique key repeats.

```{r}
# Check whether the unique key you've identified repeats
```

2. Calculate the total fatality count in this dataset and total injury count in this dataset. Calculate the percentage of NA entries in each of these variables. 

```{r}
# Calculate the total fatality count and total injury count

# Calculate the % NA values
```

3. Uncomment and complete the code below to generate a new column with a newspaper headline for each row in the dataset. Your headline should include at least five variables from the dataset, concatenated with narrative text. 

```{r}
# Generate a new column with a newspaper headline for each landslide

#global_landslide$headline <- paste()
```

4. Check the possible values in `landslide_size`. Factor this variable, setting the levels from smallest to largest. Table the unique values in `landslide_size` and `landslide_size-factored`. 

```{r}
# Check the possible values in landslide_size

# Uncomment below and factor landslide_size 
#global_landslide$landslide_size_factored <-

# Compare the outputs when you run the table() function with `landslide_size` vs. with global_landslide$landslide_size_factored

```

