---
title: "Day Five: Vector Functions and Operations"
subtitle: "SDS 192: Introduction to Data Science"
author: |
  Lindsay Poirier<br/>
  <span style = 'font-size: 70%;'>
  [Statistical & Data Sciences](http://www.smith.edu/sds), Smith College<br/>
  </span>
date: |
  Spring 2022<br/>
  `r sds::img_logo(width = 64)`
output:
  sds::moon_reader:
    lib_dir: libs
    css: xaringan-custom.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
# sds::duo_smith()
sds::mono_light_smith()
```

class: center, middle
# What were some takeaways from class last Monday?

---

# For Today

* Reviewing Large Data Frames
* Checking for row uniqueness
* Basic math operations and functions
* Basic string operations and functions
* Converting variable types
* Dealing with missing values

---

# Let's recreate the data frame from Monday to help us through today's lecture

```{r}
values <- c(1, 2, 3, 4, 5)
colors <- c("black", "red", "black", "red", "black")
suits <- c("spade", "diamond", "spade", "heart", "club")
  
# Create a new data frame for playing cards
playing_card_data <- data.frame(values, 
                                colors, 
                                suits)
```

---

# Reviewing Data Frames

* Ways to review data frames:
  * `View(playing_card_data)`
  * Click on the playing_card_data variable in your console (same as above)
  * `head(playing_card_data)`: returns first six rows of dataset
  * `names(playing_card_data)`: returns the dataset's column names
  * `nrow(playing_card_data)`: returns the number of rows in the dataset
  * `ncol(playing_card_data)`: returns the number of columns in the dataset

---

# Unique Keys

* Many datasets include a column with a number or string of characters that uniquely identify each row or observation.  
* This key should never repeat in the column, allowing us to identify each row. 
* Why might we need this?

---

# How do we make sure a column is actually functioning as a unique key?

.pull-left[
* Write some lines of code to determine whether the `values` column can act as unique key for our `playing_cards_data`.
* Hint: You need to check whether the values in the column ever repeat. 
]
.pull-right[
```{r echo=FALSE}
head(playing_card_data)
```

]

---

# What if there is no unique key?


.pull-left[
* Sometimes we use multiple columns to identify each row. 
* What two columns here would uniquely identify each row? How could we adjust our previous code to test for this?
* What if no combination of values uniquely identify each row? This means that we have duplicates in our dataset. 
]

.pull-right[
```{r echo=FALSE}
values <- c(1, 2, 3, 4, 5, 2)
colors <- c("black", "red", "black", "red", "black", "black")
suits <- c("spade", "diamond", "spade", "heart", "club", "club")
  
# Create a new data frame for playing cards
playing_card_data <- data.frame(values, 
                                colors, 
                                suits)

playing_card_data
```
]

---

# Math Operations

* `R` can function as a calculator:

```{r}
3 + 3
8 / 2
2 * 2
```

---

# Math Operations

* Why does this produce an error?

```{r, error=TRUE}
"3" + "3"
```

---

# Math Functions on Vectors

* Let's say I wanted to calculate the sum of my hand of cards. We can pass the `values` column as an argument to the `sum()` function. 

```{r}
sum(playing_card_data$values)
```

* What if I wanted to calculate the maximum value in my hand?
* What if I wanted to calculate the median value in my hand?

---

# Math Functions on Vectors

* Why does this produce an error? 

```{r, error=TRUE}
sum(playing_card_data$colors)
```

---

# String Operations

* We can *concatenate* strings by taking two separate strings and combining them into a single string. We use the `paste()` function for this. 
* Enter `?paste()` into the console. What arguments are required?

---

# String Operations

* We can *concatenate* strings by taking two separate strings and combining them into a single string. We use the `paste()` function for this. 

```{r}
word1 <- "Harry"
word2 <- "Sally"
paste("When", word1, "Met", word2, sep = " ")
```

---

# String Operations

* Why doesn't this produce an error?

```{r, error=TRUE}
word1 <- 2.3
word2 <- 34
paste("When", word1, "Met", word2, sep = " ")
```

---

# String Functions on Vectors

* Let's say I wanted to convert all of the characters in the `colors` column to uppercase. We can pass the `colors` column to the `toupper()` function. 

```{r}
toupper(playing_card_data$colors)
```

* What if I wanted to convert the characters to lowercase?

---

# Converting Variable Types

* What if we wanted to produce a sentence that listed the values of the cards in our hand?

```{r}
playing_card_data$string_values <- 
  as.character(playing_card_data$values)

part1 <- paste(playing_card_data$string_values, 
               collapse = ", ")
paste("I'm holding a", part1, sep = " ")
```

> Challenge: How might we add the word "and" between the last two values?

---

# Converting Variable Types

* How might we add the word "and" between the last two values??

```{r}
playing_card_data$string_values <- 
  as.character(playing_card_data$values)

part1 <- paste(playing_card_data$string_values[1:5], 
               collapse = ", ")
part2 <- playing_card_data$string_values[6]
paste("I'm holding a", part1, "and", part2, sep = " ")
```

> At-home Challenge: How might we add an Oxford comma?

---

# Converting Variable Types

* What if we were supplied these numbers as characters?

```{r, error=TRUE}
playing_card_data$values <- 
  c("1", "2", "3", "4", "5", "6")
sum(playing_card_data$values)
```

```{r}
playing_card_data$values <- 
  as.numeric(playing_card_data$values)
sum(playing_card_data$values)
```

---

# Factors

* What if we were dealt a new hand of cards that only included face cards? What kind of variable is this? How do we indicate the order of values to R?

```{r}
playing_card_data$values <- 
  c("K", "Q", "J", "Q", "A", "K")
sort(playing_card_data$values)

playing_card_data$values <- factor(playing_card_data$values, 
                                   levels = c("J", "Q", "K", "A"))
sort(playing_card_data$values)
```

---

# Missing Values

* Remember that missing values still have a position in rectangular datasets
* Missing values get recorded as `NA` in R
* ...but sometimes analysts put words or numbers in their datasets to indicate missingness:
  * "NONE"
  * -999
  * "" <- this is the most challenging to uncover!
* In a few weeks, we will talk about how to convert these to NAs
* ...but what happens when we try to perform functions on vectors that contain missing values?

---

# Missing Values in Math Functions

* We can use `na.rm = TRUE` to ignore NA values in math functions. 

```{r}
playing_card_data$values <- c(1, 2, NA, 4, NA, 6)

sum(playing_card_data$values)
sum(playing_card_data$values, na.rm = TRUE)
```

---

# Checking number of missing values

* We can use `is.na()` to check if a value is an NA value. This will return TRUE or FALSE.
* What other function might we use to check the number of NA values in a vector?

```{r}
playing_card_data$values <- c(1, 2, NA, 4, NA, 6)

sum(is.na(playing_card_data$values))
```

---

# For Wednesday

* No reading. Keep working through problem sets!
* Wednesday: Applying functions to data frame values
* Friday: Learning to code: debugging, cheatsheets, etc. 

